Proposal outline/notes for DPG and Compressible Navier-Stokes

### 2012-06-19 docs/Jesse/Proposal/DPGtheory.tex
adding more additions (NL DPG) to proposal
committing changes to proposal tex/bib files

The Discontinuous Petrov-Galerkin (DPG) method of Demkowicz and Gopalakrishnan was first formulated in \cite{DPG1} for the pure convection problem.

Historically, the name DPG was given by Bottasso, Micheletti, Sacco and Causin to their method for elliptic problems in \cite{BottassoMichelettiSacco02}. The method was then extended to other problems, including convection-diffusion, in \cite{BottassoMichelettiSacco05,CausinSacco05,CausinSaccoBottasso05}.

The key point in the method of Bottasso et al.\ was their method of hybridization of fluxes. Whereas HDG methods will identify an additional flux unknown on the boundary, the numerical flux coupling neighboring elements together is still typically computed in part using contributions from the field variables. In the DPG formulation, all numerical fluxes are declared to be independent unknowns, leaving the interior field degrees of freedom completely uncoupled from element to element. \todo{Finish, check if this is true.}

The breakthrough came in \cite{DPG2}, where the concept of locally-computable optimal test functions led to the development of the DPG method in its current form.

\todo{Clean up}
This DPG method with optimal test functions is better understood in \cite{DPG3,DPG4}, extended to problems in \cite{DPGElas,stokesDPG}, well-posedness theory developed for Poisson and convection-diffusion in \cite{analysisDPG}, extended to large class of Friedrichs systems in \cite{Bui-ThanhDemkowiczGhattas11b}. \cite{practicalDPG}.

Optimal order comparison to DG in \cite{DPG1}. Comparisons to other DG methods in \cite{Bui-ThanhDemkowiczGhattas11a,DGDPG}.

\section{Optimal Petrov-Galerkin methods}

@@

\section{Ultra-weak variational formulation}
\seclab{abstractUweak}

@@

\section{Choices of test and trial norms}
\seclab{energyPair}

@@

Unfortunately, the optimal test norm is non-localizable due to the presence of the jump term $\jump{v}$.\footnote{A localizable norm $\nor{v}_{V(\Oh)}$ can be written in the form $\nor{v}_{V(\Oh)} = \sum_{K\in\Oh} \nor{v}_{V(K)},$ where $\nor{v}_{V(K)}$ is a norm over the element $K$.} Since the jump terms couple elements together, the evaluation of the jump terms requires contributions from all the elements in the mesh. Consequently, solving for an optimal test function
amounts to inverting the Riesz map %, defined on the left sideof \eqnref{optvVar}, 
over the entire mesh $\Oh$, making the optimal test norm impractical.

>>>>

The Discontinuous Petrov-Galerkin (DPG) method of Demkowicz and Gopalakrishnan was first formulated in \cite{DPG1} as a scheme for the pure convection problem. The method demonstrated promise --- in particular, the method was able to achieve optimal convergence rates for the Peterson problem where standard DG methods were suboptimal. A breakthrough came in \cite{DPG2}, where the concept of locally-computable optimal test functions led to the development of the DPG method in its current form. Soon after, the DPG method was successfully applied to solve the convection-diffusion problem in the small-diffusion limit with $\epsilon = 1e-7$ \cite{DPG2,DPG3}.

Historically, the name DPG was given by Bottasso, Micheletti, Sacco and Causin to their method for elliptic problems in \cite{BottassoMichelettiSacco02}. The method was then extended to other problems, including convection-diffusion, in \cite{BottassoMichelettiSacco05,CausinSacco05,CausinSaccoBottasso05}. The key point in the method of Bottasso et al.\ was their method of hybridization of fluxes. Whereas HDG methods identify an additional flux unknown on the boundary, the numerical flux coupling neighboring elements together is still typically computed in part by using contributions from interior degrees of freedom on neighboring elements. In the DPG formulation, all numerical fluxes are declared to be independent unknowns, leaving the interior field degrees of freedom completely uncoupled from element to element. This formulation is often referred to as the \emph{ultra-weak} variational formulation.

Since its inception, the DPG method has been applied to a wide range of problems with great success, including elasticity \cite{DPGElasLocking}, thin body shell problems \cite{DPGElas}, the cloacking problem in electromagnetics \cite{DPGcloak}, both Helmholtz and elastic wave propagation problems \cite{DPG4}, and recently, the linear Stokes equation \cite{stokesDPG}. DPG has also been shown to be a generalization of several successful finite element methods --- comparisons between DPG and standard DG methods can be found in \cite{DPG1}, and more recently, several existing DG methods have been shown to be derivable using the DPG framework \cite{Bui-ThanhDemkowiczGhattas11a,DGDPG}.

Detailed analysis of the energy setting and well-posedness of DPG has been done for the Poisson and convection-diffusion equations in \cite{analysisDPG}. The well-posedness of DPG has also recently been extended to the large class of Friedrichs systems in \cite{Bui-ThanhDemkowiczGhattas11b}. Significant efforts have also been made in demonstrating the robustness of the DPG method for singular perturbation problems, both in wave propagation \cite{DPG4,DPGwave} and convection-diffusion problems \cite{DPGrobustness, DPGrobustness2}.

\subsection{Optimal Petrov-Galerkin methods}

@@

\subsection{Ultra-weak variational formulation}
\seclab{abstractUweak}

@@

\subsection{Choices of test and trial norms}
\seclab{energyPair}

@@

Unfortunately, the optimal test norm is non-localizable due to the presence of the jump term $\jump{v}$.\footnote{A localizable norm can be written as $\nor{v}_{V(\Oh)} = \sum_{K\in\Oh} \nor{v}_{V(K)},$ where $\nor{v}_{V(K)}$ is a norm over $K$.} Since the jump terms couple elements together, the evaluation of the jump terms requires contributions from all the elements in the mesh. Consequently, solving for an optimal test function
amounts to inverting the Riesz map %, defined on the left sideof \eqnref{optvVar}, 
over the entire mesh $\Oh$, making the optimal test norm impractical.

### 2012-06-19 docs/Jesse/Proposal/NSNotes.tex
committing changes to proposal tex/bib files

Following Demkowicz and Oden \cite{Demkowicz:1990:NFE:112271.112276}, we consider the compressible Navier-Stokes equations
\[
\sum^2_{i=1} F_i\left( U\right)_{,i}=\sum^2_{i=1} G_i\left( U\right)_{,i}
\]
where the Eulerian fluxes are given as
\[
F_1 = \vectfour{\rho u_1}{\rho u_1^2+p}{\rho u_1 u_2}{((\rho e)+p)u_1}, \qquad F_2 = \vectfour{\rho u_2}{\rho u_1 u_2}{\rho u_2^2+p}{((\rho e)+p)u_2},
\]
and the viscous fluxes are given as
\[
G_i = \vectfour{0}{\sigma_{i1}}{\sigma_{i2}}{\sum_{j=1}^2\sigma_{ij}u_j + q_i}.
\]
We have also the constitutive relations 
\begin{align*}
\sigma_{ij} &= \mu(u_{i,j}+u_{j,i}) + \lambda u_{k,k}\delta_{ij}\\
q_i &= \kappa T_{,i}\\
p &= (\gamma-1)\rho\iota\\
\iota &= e-\frac{1}{2}(u_1^2+u_2^2)\\
\iota &= c_vT
\end{align*}
where $e$ and $\iota$ are energy and internal energy per unit mass, respectively. We write the Navier Stokes equations as a system of seven first order equations. Let $[u_1,u_2] = [u,v]$.

\begin{itemize}
\item \textbf{Mass equation}
\[
\div \vecttwo{\rho u }{\rho v} = 0
\]
\item \textbf{Momentum equations}
\begin{align*}
\div \vecttwo{\rho u^2+p }{\rho u v} &= \div\left(\vec{\sigma_{i1}}\right)\\
\div \vecttwo{\rho u v}{\rho v^2+p } &= \div\left(\vec{\sigma_{i2}}\right)
\end{align*}

\item \textbf{Energy equation}

Let $\boldsymbol \sigma$ be a tensor whose $ij$th term is $\sigma_{ij}$.  
\[
\div\vecttwo{((\rho e)+p)u}{((\rho e)+p)v} = \div\left[\boldsymbol\sigma U + \vec{q}\right]
\]

@@

We represent $\boldsymbol\sigma$ using the Newtonian fluid law
\[
\sigma_{ij} = \mu(u_{i,j} + u_{j,i}) + \lambda u_{k,k} \delta_{ij}
\]
where $\mu$ is viscosity and $\lambda$ is bulk viscosity. 
%Following elasticity, a general formula for the stress tensor is 
%\[
%\sigma_{ij} = 2\mu \epsilon_{ij} + \lambda \epsilon_{kk} \delta_{ij}
%\]
We can invert the stress tensor under isotropic and plane strain assumptions to get
%\[
%\epsilon_{ij} = \frac{1}{2\mu} \sigma_{ij} - \frac{\lambda}{4\mu (\mu + \lambda)} \sigma_{kk}\delta_{ij}
%\] 
%which translates to
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \frac{1}{2\mu} \sigma_{ij} - \frac{\lambda}{4\mu (\mu + \lambda)} \sigma_{kk}\delta_{ij}
\]
We also have
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \grad  U - \boldsymbol \omega
\]
where $\boldsymbol \omega$ is the antisymmetric part of the infinitesimal strain tensor:
\[
\boldsymbol \omega = \frac{1}{2}\left(\grad  U - \grad ^T  U\right).
\]
Thus our final form is
\begin{align*}
\grad  U - \boldsymbol \omega = \frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I.
\end{align*}
Notice that $\boldsymbol \omega$ is implicitly defined to be the symmetric part of $\grad u$ by taking the symmetric part of the above equation.

@@

In general, $\mu$ and $\lambda$ are functions of temperature, obeying the power law
\[
\mu = \left(\frac{T}{T_0}\right)^\beta,
\]
where $T_0$ is a reference temperature. We choose $\beta = 2/3$ in this case.

@@

We assume Fourier's law 
\begin{align*}
\vec{q} &= \kappa \grad T,
\end{align*}
where $\kappa$ is generally a function of temperature. We introduce here the Prandtl number here as well
\[
{\rm Pr} = \frac{\gamma c_v \mu}{\kappa}.
\]
In this case, we assume a constant Prandlt number, which implies that the heat conductivity $\kappa$ is proportional to viscosity $\mu$.

@@

\subsection{Nondimensionalization}
To nondimensionalize our equations, we introduce nondimensional quantities for length, density, velocity, temperature, and viscosity. 
\[
\boldsymbol x^* = \frac{\boldsymbol x}{L}, \qquad \rho^* = \frac{\rho}{\rho_{\infty}}, \qquad \boldsymbol u^* = \frac{\boldsymbol u}{V_\infty}, \qquad T^* = \frac{T}{T_\infty}, \qquad \mu^* = \frac{\mu}{\mu_\infty}
\]
Momentum, pressure, internal energy, and bulk viscosity are then nondimensionalized with respect to the above variables
\[
\boldsymbol m^* = \frac{\boldsymbol m}{\rho_\infty V_\infty}, \qquad p^* = \frac{p}{\rho_\infty V_\infty^2}, \qquad \iota^* = \frac{\iota}{V_\infty^2}, \qquad \lambda^* = \frac{\lambda}{\mu_\infty}
\]
We introduce, for convenience, the Reynolds number
\[
\Reyn = \frac{\rho_\infty V_\infty L}{\mu_\infty} 
\]
and the reference (free stream) Mach number
\[
M_\infty = \frac{V_\infty}{\sqrt{\gamma(\gamma-1)c_vT_\infty}}
\]
Note that 
\[
a = \sqrt{\frac{\gamma p_\infty}{\rho_\infty}} = \sqrt{{\gamma p_\infty}} = \sqrt{\gamma(\gamma-1)c_vT_\infty}
\]
The equations take the same form as before after nondimensionalization, so long as we define new material constants
\[
\tilde{\mu} = \frac{\mu^*}{\Reyn}, \qquad \tilde{\lambda} = \frac{\lambda^*}{\Reyn}, \qquad \tilde{c}_v = \frac{1}{\gamma(\gamma-1)M_\infty^2}, \qquad \tilde{\kappa} = \frac{\gamma\tilde{c}_v\tilde{\mu}}{\rm Pr}
\]
From here on, we drop the $^*$ superscript and assume all variables refer to their nondimensionalized quantities.

@@

The Navier-Stokes conservation laws can be written as 
\begin{align*}
\div \vecttwo{\rho u }{\rho v} &= 0\\
\div \left(\vecttwo{\rho u^2+p }{\rho u v} - \boldsymbol \sigma_{1}\right) &=0\\
\div \left(\vecttwo{\rho u v}{\rho v^2+p } - \boldsymbol \sigma_{2}\right) &=0\\
\div \left(\vecttwo{((\rho e)+p)u}{((\rho e)+p)v} - \boldsymbol \sigma_1 \cdot \boldsymbol u- \boldsymbol \sigma_2 \cdot \boldsymbol u + \vec{q}\right) &=0\\
\end{align*}
or generally, 
\[
\div (F_i(\boldsymbol U)-G_i(\boldsymbol U,\boldsymbol \sigma) = 0, \qquad i = 1,\ldots, 4
\]
The variational form restricted to a single element gives
\[
\langle \widehat{F}_i\cdot n, v\rangle - \int_K  (F(\boldsymbol U)-G_i(\boldsymbol U,\boldsymbol \sigma)) \cdot \grad v_i = 0 , \qquad i = 1,\ldots, 4
\]
and the variational form over the entire domain is given by summing up the element-wise contributions.

@@

\subsection{Numerical experiments}

For our solver, we use a standard pseudo-time solver and a standard greedy refinement scheme. Though we have not yet implemented adaptive timestepping, we are able to take large enough uniform timesteps over a coarse mesh such that convergence is comparable to the convergence of a damped Newton method.

The numerical parameters used are as follows:
\begin{itemize}
\item{DPG parameters:} Both $p$ and $\Delta p = 2$ uniformly across the mesh. 
\item{Adaptivity parameters:} Energy threshold for refinements is $\alpha = .2$ or $\alpha = .15$. 
\item{Time-stepping parameters:} $\Delta t = .1$, and tolerance for transient residual $\epsilon_t = 5e-7$.
\end{itemize}

\subsubsection{Carter flat plate}

Our first problem of interest is the Carter flat plate problem. An infinitesmally thin flat plate disrupts a free stream flow and causes a shock to form at the tip of the plate.

@@

\begin{figure}[!h]
\centering
\subfigure[$\rho$]{\includegraphics[scale=.42]{figs/rho.pdf}}
\subfigure[$u_1$]{\includegraphics[scale=.42]{figs/u1.pdf}}
\subfigure[$u_2$]{\includegraphics[scale=.42]{figs/u2.pdf}}
\subfigure[$T$]{\includegraphics[scale=.42]{figs/T.pdf}}
\caption{Solutions after six refinements for $\Reyn = 1000$.}
\label{fig:Re500}
\end{figure}

\begin{figure}[!h]
\centering
\subfigure[$\rho$]{\includegraphics[scale=.4]{figs/rhozoom.pdf}}
\subfigure[$u_1$]{\includegraphics[scale=.4]{figs/u1zoom.pdf}}
\subfigure[$u_2$]{\includegraphics[scale=.4]{figs/u2zoom.pdf}}
\subfigure[$T$]{\includegraphics[scale=.4]{figs/Tzoom.pdf}}
\caption{Zoom of solutions at the beginning of the plate for $\Reyn = 1000$.}
\label{fig:Re500}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[scale=.75]{figs/mesh.pdf}
\caption{Adaptive mesh after six refinements for $\Reyn = 1000$.}
\label{fig:MeshRe500}
\end{figure}

>>>>

We briefly review the compressible Navier-Stokes equations, given in Section~\secref{sec:compNS}, and formulate DPG for the nonlinear system. 
\begin{itemize}
\item \textbf{Conservation equations}
\begin{align*}
\div \vecttwo{\rho u_1 }{\rho u_2} &= 0\\
\div \vecttwo{\rho u_1^2+p }{\rho u_1 u_2} &= \div\left(\vec{\sigma_{i1}}\right)\\
\div \vecttwo{\rho u_1 u_2}{\rho u_2^2+p } &= \div\left(\vec{\sigma_{i2}}\right)\\
\div\vecttwo{((\rho e)+p)u_1}{((\rho e)+p)u_2} &= \div\left[\boldsymbol\sigma U + \vec{q}\right],
\end{align*}
where $\boldsymbol \sigma$ is the stress tensor whose $ij$th term is $\sigma_{ij}$.

@@

We represent $\boldsymbol\sigma$ using the Newtonian fluid law
\[
\sigma_{ij} = \mu(u_{i,j} + u_{j,i}) + \lambda u_{k,k} \delta_{ij}
\]
where $\mu$ is viscosity and $\lambda$ is bulk viscosity. 
%Following elasticity, a general formula for the stress tensor is 
%\[
%\sigma_{ij} = 2\mu \epsilon_{ij} + \lambda \epsilon_{kk} \delta_{ij}
%\]
We can invert the stress tensor under isotropic and plane strain assumptions to get
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \frac{1}{2\mu} \sigma_{ij} - \frac{\lambda}{4\mu (\mu + \lambda)} \sigma_{kk}\delta_{ij}
\]
We also have
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \grad  U - \boldsymbol \omega
\]
where $\boldsymbol \omega$ is the antisymmetric part of the infinitesimal strain tensor:
\[
\boldsymbol \omega = \frac{1}{2}\left(\grad  U - \grad ^T  U\right).
\]
Thus our final form is
\begin{align*}
\grad  U - \boldsymbol \omega = \frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I.
\end{align*}
Notice that $\boldsymbol \omega$ is implicitly defined to be the symmetric part of $\grad u$ by taking the symmetric part of the above equation.

@@

@@

We assume Fourier's law 
\begin{align*}
\vec{q} &= \kappa \grad T,
\end{align*}
%where $\kappa$ is generally a function of temperature. 
We introduce here the Prandtl number here as well
\[
{\rm Pr} = \frac{\gamma c_v \mu}{\kappa}.
\]
In this case, we assume a constant Prandlt number, which implies that the heat conductivity $\kappa$ is proportional to viscosity $\mu$.

@@

\subsection{Nondimensionalization}
To nondimensionalize our equations, we introduce nondimensional quantities for length, density, velocity, temperature, and viscosity. 
\[
\boldsymbol x^* = \frac{\boldsymbol x}{L}, \qquad \rho^* = \frac{\rho}{\rho_{\infty}}, \qquad u_1^* = \frac{u_1}{V_\infty}, \qquad u_2^* = \frac{u_2}{V_\infty}, \qquad T^* = \frac{T}{T_\infty}, \qquad \mu^* = \frac{\mu}{\mu_\infty}
\]
Pressure, internal energy, and bulk viscosity are then nondimensionalized with respect to the above variables
\[
p^* = \frac{p}{\rho_\infty V_\infty^2}, \qquad \iota^* = \frac{\iota}{V_\infty^2}, \qquad \lambda^* = \frac{\lambda}{\mu_\infty}
\]
We introduce, for convenience, the Reynolds number
\[
\Reyn = \frac{\rho_\infty V_\infty L}{\mu_\infty} 
\]
and the reference (free stream) Mach number
\[
M_\infty = \frac{V_\infty}{\sqrt{\gamma(\gamma-1)c_vT_\infty}}
\]
Note that 
\[
a = \sqrt{\frac{\gamma p_\infty}{\rho_\infty}} = \sqrt{{\gamma p_\infty}} = \sqrt{\gamma(\gamma-1)c_vT_\infty}
\]
The equations take the same form as before after nondimensionalization, so long as we define new material constants
\[
\tilde{\mu} = \frac{\mu^*}{\Reyn}, \qquad \tilde{\lambda} = \frac{\lambda^*}{\Reyn}, \qquad \tilde{c}_v = \frac{1}{\gamma(\gamma-1)M_\infty^2}, \qquad \tilde{\kappa} = \frac{\gamma\tilde{c}_v\tilde{\mu}}{\rm Pr}
\]
From here on, we drop the $^*$ superscript and assume all variables refer to their nondimensionalized quantities.

@@

The Navier-Stokes conservation laws can be written as 
\begin{align*}
\div \vecttwo{\rho u }{\rho v} &= 0\\
\div \left(\vecttwo{\rho u^2+p }{\rho u v} - \boldsymbol \sigma_{1}\right) &=0\\
\div \left(\vecttwo{\rho u v}{\rho v^2+p } - \boldsymbol \sigma_{2}\right) &=0\\
\div \left(\vecttwo{((\rho e)+p)u}{((\rho e)+p)v} - \boldsymbol \sigma_1 \cdot \boldsymbol u- \boldsymbol \sigma_2 \cdot \boldsymbol u + \vec{q}\right) &=0
\end{align*}
or generally, 
\[
\div (F_i(\boldsymbol U)-G_i(\boldsymbol U,\boldsymbol \sigma) = 0, \qquad i = 1,\ldots, 4
\]
The variational form restricted to a single element gives
\[
\langle \widehat{F}_i\cdot n, v\rangle - \int_K  (F(\boldsymbol U)-G_i(\boldsymbol U,\boldsymbol \sigma)) \cdot \grad v_i = 0 , \qquad i = 1,\ldots, 4
\]
and the variational form over the entire domain is given by summing up the element-wise contributions.

@@

\subsection{Numerical experiments: Carter flat plate}

For our solver, we use a standard pseudo-time solver and greedy refinement scheme. Though we have not yet implemented adaptive timestepping, we are able to take large enough uniform timesteps over a coarse mesh such that convergence is order-of-magnitude comparable to the convergence of a damped Newton method.

The numerical parameters used are as follows:
\begin{itemize}
\item{DPG parameters:} $p = 2$ or $p=3$, and $\Delta p = 2$ uniformly across the mesh. 
\item{Adaptivity parameters:} Energy threshold for refinements is $\alpha = .2$ or $\alpha = .15$. 
\item{Time-stepping parameters:} $\Delta t = .1$, and tolerance for transient residual $\epsilon_t = 5e-7$.
\end{itemize}

Our problem of interest is the Carter flat plate problem. An infinitesmally thin flat plate disrupts a free stream flow and causes a shock to form at the tip of the plate.

@@

\begin{figure}[!h]
\centering
\subfigure[$\rho$]{\includegraphics[scale=.42]{figs/Re1000p2/rho.pdf}}
\subfigure[$u_1$]{\includegraphics[scale=.42]{figs/Re1000p2/u1.pdf}}
\subfigure[$u_2$]{\includegraphics[scale=.42]{figs/Re1000p2/u2.pdf}}
\subfigure[$T$]{\includegraphics[scale=.42]{figs/Re1000p2/T.pdf}}
\caption{Solutions after six refinements for $p=2$ and $\Reyn = 1000$.}
\label{fig:Re500}
\end{figure}

\begin{figure}[!h]
\centering
\subfigure[$\rho$]{\includegraphics[scale=.4]{figs/Re1000p2/rhozoom.pdf}}
\subfigure[$u_1$]{\includegraphics[scale=.4]{figs/Re1000p2/u1zoom.pdf}}
\subfigure[$u_2$]{\includegraphics[scale=.4]{figs/Re1000p2/u2zoom.pdf}}
\subfigure[$T$]{\includegraphics[scale=.4]{figs/Re1000p2/Tzoom.pdf}}
\caption{Zoom of solutions at the beginning of the plate for $p=2$ and $\Reyn = 1000$.}
\label{fig:Re500}
\end{figure}

For $p=2$, after six refinements, the mesh contained 1410 elements and 187605 global dofs, with a total DPG energy error of $5.684e-7$. Comparatively, for $p=3$, the mesh contained only 444 elements and 97403 global dofs, with a total DPG eneryg error of 4.57681e-7 after five refinements.

\begin{figure}[!h]
\centering
\subfigure[$p=2$]{\includegraphics[scale=.5]{figs/Re1000p2/mesh.pdf}}
\subfigure[$p=3$]{\includegraphics[scale=.5]{figs/Re1000p3/mesh.pdf}}
\caption{Adaptive mesh after six refinements for $p=2$ and five refinements for $p=3$ with $\Reyn = 1000$.}
\label{fig:MeshRe500}
\end{figure}

### 2012-06-19 docs/Jesse/Proposal/NonlinDPG.tex
committing changes to proposal tex/bib files

In practice, most compressible flow solvers opt for the pseudo-time step method over the direct Newton iteration due to the difficulty of convergence and sensitivity of the Newton iteration to initial guess\cite{libMeshPaper}.  Though the convergence of the pseudo-time step is slower, the addition of the zero-order transient terms ``regularizes" the problem, making it easier to solve.\footnote{The addition of a zero-order term ``regularizes" an equation by adding to it a positive-definite $L^2$ projection operator.  In the limit as $\Delta t\rightarrow 0$, the solution at $t_i$ will simply return the $L^2$ projection of the solution at the previous timestep.}

@@

\todo{Add adjoint trick stuff}

>>>>

In practice, most compressible flow solvers opt for the pseudo-time step method over the direct Newton iteration due to the difficulty of convergence and sensitivity of the Newton iteration to initial guess\cite{libMeshPaper}.  Though the convergence of the pseudo-time step is slower, the addition of the zero-order transient terms ``regularizes" the problem and makes it less difficult to solve.\footnote{The addition of a zero-order term ``regularizes" an equation by adding to it a positive-definite $L^2$ projection operator.  In the limit as $\Delta t\rightarrow 0$, the solution at $t_i$ will simply return the $L^2$ projection of the solution at the previous timestep.}

@@

A recent theoretical development is the formulation of a DPG method that aims to minimize a nonlinear residual. Given two Hilbert spaces --- a trial space $U$ and test space $V$ --- our nonlinear variational formulation can be written as $b(u,v) = l(v)$, with the corresponding operator form of the formulation in $V'$
\[
B(u) = l.
\]
We can apply the steps used to derive the DPG method for linear problems to the nonlinear setting as well. Given a finite dimensional subspace $U_h \subset U$, we consider the discrete nonlinear residual
\[
J(u_h) \coloneqq \frac{1}{2}\nor{R_V^{-1}\left(B(u_h)-l\right)}_V^2.
\]
Our goal is to solve
\[
u_h = \arg \min_{w_h\in U_h}J(w_h).
\]
Similarly to the linear case, we can take the Gate\'aux derivative to arrive at a necessary condition for $u_h$ to minimize $J(u_h)$ 
\begin{align*}
\LRa{J'(u_h), \delta u_h} &= \LRp{R_V^{-1}\left(B(u_h)-l\right),R_V^{-1}B'(u_h;\delta u_h)}_V, \quad \delta u_h \in U_h. 
\end{align*}
As the above is a nonlinear equation, we seek its solution through linearization. Differentiating a second time in $u$, we arrive at
\begin{align*}
\LRa{J''(u_h), \Delta u_h} =& \LRa{B'(u_h;\Delta u_h), B'(u_h;\delta u_h)}_V \\
& + \LRa{\left(B(u_h)-l\right),B''(u_h;\delta u_h,\Delta u_h)}_V \\
=& \LRp{R_V^{-1}B'(u_h;\Delta u_h), R_V^{-1}B'(u_h;\delta u_h)}_V \\
& + \LRp{R_V^{-1}\left(B(u_h)-l\right),R_V^{-1}B''(u_h;\delta u_h,\Delta u_h)}_V 
\end{align*}
where $B''(u_h;\delta u_h,\Delta u_h)$ denotes the Hessian of $B(u_h)$, evaluated using both $\delta u_h$ and $\Delta u_h$.

Examining the above formulation, we note that DPG as applied to the linearized problem produces the term $ \LRp{R_V^{-1}B'(u_h;\Delta u_h), R_V^{-1}B'(u_h;\delta u_h)}_V$. However, in approaching the nonlinear problem through the minimization of the discrete residual, we gain a second-order term involving the Hessian
\[
\LRp{R_V^{-1}\left(B(u_h)-l\right),R_V^{-1}B''(u_h;\delta u_h,\Delta u_h)}_V.
\] 
The evaluation of this term can be done in a computationally efficient manner --- if we define the image of the nonlinear residual under the Riesz inverse
\[
v_{R(u)} = R_V^{-1}\left(B(u_h)-l\right),
\]
then we can compute this additional term through
\begin{align*}
\LRp{v_{R(u)},R_V^{-1}B''(u_h;\delta u_h,\Delta u_h)}_V &= \LRa{v_{R(u)},B''(u_h;\delta u_h,\Delta u_h)}_V
\end{align*}
which can be computed in the same fashion as a Bubnov-Galerkin stiffness matrix. This addition, though not positive definite, is symmetric due to the nature of second order Ga\^teux derivatives.

We note that we have not implemented this Hessian-based nonlinear solver in the numerical experiments to follow, and instead plan to do so in the proposed work outlined in Section~\secref{sec:proposedWork}.

### 2012-06-19 docs/Jesse/Proposal/areaReq.tex
committing changes to proposal tex/bib files

\section{Area requirements}

@@

\begin{itemize}
\item{\textbf{Completed: }} Prove robustness of DPG method for the scalar convection-diffusion problem
\item Attempt analysis of the linearized Navier-Stokes system
\end{itemize}

@@

\begin{itemize}
\item{\textbf{Completed: }} Collaborative work with Nathan Roberts on the higher order parallel adaptive DPG code Camellia.
\item Nonlinear DPG - Hessian adjoint trick
\item Adaptivity: anisotropic refinements and $hp$-schemes.
\item Distributed static condensation
\end{itemize}

@@

\begin{itemize}
\item{\textbf{Completed: }} Convection-dominated diffusion, Burgers, and Navier-Stokes. 
\item Higher Reynolds number, ramp problem, Gaussian bump, airfoil.
\item Euler with NS regularization?
\end{itemize}

>>>>

\section{Area requirements}
\seclab{sec:proposedWork}

@@

\begin{itemize}
\item{\textbf{Completed: Prove robustness of DPG method for the scalar convection-diffusion problem.}}

Analysis has been done in \cite{DPGrobustness2}, extending the results of \cite{DPGrobustness}. In particular, we have introduced a test norm under which the DPG method robustly bounds the $L^2$ error in the field variables $u$ and the scaled stress $\sigma$. Numerical results confirm the theoretical bounds given.

\item{\textbf{Proposed: Attempt analysis of the linearized Navier-Stokes system.}}

We hope to analyze the linearized Navier-Stokes equations to determine an optimal extrapolation of the test norm for the scalar convection-diffusion problem to systems.

\end{itemize}

@@

\begin{itemize}
\item{\textbf{Completed: Collaborative work with Nathan Roberts on the higher order parallel adaptive DPG code Camellia.}}

Most numerical experiments have been done using the higher-order adaptive codebase Camellia, built upon the Trilinos library and designed by Nathan Roberts. The library allows for rapid implementation of problems through a symbolic syntax similar to the Fenics project. Work has been done to support arbitrary-irregularity refinements in both $h$ and $p$, and the framework for anisotropic refinements is in place as well. Similarly, the code is partially parallelized - the determination of the stiffness matrix is perfectly scalable, and we are currently exploring options for more scalable solvers.

\item{\textbf{Proposed: Distributed iterative static condensation.}}

A clear choice for a parallelized solver under the ultra-weak variational formulation is static condensation/the Schur-complement method. Given a block matrix structure of a stiffness matrix $K$, we can view the DPG system as
\[
Ku = \arr{A}{B}{B^T}{D}\vecttwo{u_{\rm flux}}{u_{\rm field}} = \vecttwo{f}{g} = l
\]
where $D$ has a block-diagonal structure, and $A$ and $D$ are both square matrices with $\dim{A} < \dim{D}$. The system can be reduced to yield the condensed system
\[
\left(A-B D^{-1} B^T\right)u_{\rm flux} = f - B D^{-1} g
\]
where $D^{-1}$ can be inverted block-wise. Once the globally coupled flux degrees of freedom are solved for, the field degrees of freedom can be reconstructed locally.

Additionally, though the Schur complement $A-B D^{-1} B^T$ is generally a dense matrix, it does not have to be explicitly formed, and we can use a matrix-free method to solve the above condensed system. The question remains on whether or not the condensed DPG stiffness matrix is well-conditioned. It has been shown that, unlike standard least-squares methods, DPG generates for the Poisson matrix a stiffness matrix with condition number $O(h^{-2})$ \cite{practicalDPG}. It is well known that, under standard finite element methods, if the condition number of the global stiffness matrix $K$ is $O(h^{-2})$, the condition number of the Schur complement is $O(h^{-1})$. Additionally, through either diagonal preconditioning or matrix equilibration, the condition number of the Schur complement can often be made significantly smaller than $O(h^{-1})$.

We hope to implement such a distributed solver and determine the scalability of this solver under DPG.

\item{\textbf{Proposed: a Nonlinear Hessian-based DPG method.}}

We hope to implement the Hessian-based nonlinear DPG approach to the viscous Burgers and compressible Navier-Stokes problems and compare convergence and error rates for each problem after applying this nonlinear strategy.

\item{\textbf{Proposed: Anisotropic refinements and $hp$-schemes.}}

The effectiveness and necessity of anisotropic refinement schemes for problems in CFD has been demonstrated several times in the literature \cite{anisotropy1,anisotropy2}. As the error representation function has been shown to yield an effective and natural vector-valued residual with which to drive refinement, we hope to generalize the use of the error representation function to yield anisotropic adaptive schemes. Additionally, we hope to explore the possibility of inferring an optimal choice between $h$ and $p$ refinement using the error representation function. %\cite{BoeingHigherOrder}.

\end{itemize}

@@

\begin{itemize}
\item{\textbf{Completed: convection-dominated diffusion, Burgers, and a model problem for Navier-Stokes.}} 
\item{\textbf{Proposed: Higher Reynolds number, ramp problem, Gaussian bump, airfoil.}}
\item{\textbf{Proposed: regularized Euler.}}
\end{itemize}
